Model Search in Relational Domains

### 2016-10-22 introduction.tex
line breaks
intro

\section{Introduction}
%
%%%%%%%%%%%%%%%%%%%%%%% AI %%%%%%%%%%%%%%%%%%%%%%%%
Static program analysis based on abstract 
interpretation~\cite{CC77} has been widely used to
verify certain classes of properties for safety-critical systems.  A 
standard static analyses~\cite{se2011} for program safety aims to compute 
invariants as fixed-points of abstract transformers.  Most lattice structure 
used in static analysis have meet and join operations.  The meet operation 
precisely model conjunction, and the join operation over-approximate 
disjunction. Thus, the precision loss in static analysers is often 
due to imprecise join operations. The loss in precision is overcome 
by using richer abstract domains or equipping the analysis with disjunction --
which is often very expensive since the analysis exhibit case enumeration 
behaviour. 
%%%%%%%%%%%%%%%%%%%%%%% ACDCL %%%%%%%%%%%%%%%%%%%%%
On the other hand, the combination of precise conjunction in a partial
assignments domain with learning allows Conflict Driven Clause
Learning (CDCL)~\cite{cdcl} based satisfiability (SAT) solvers to
reason about disjunction without enumerating cases.  Silva
et. al.~\cite{tacas12, sas12, dhk2013-popl} presented a formal recipe
for applying abstract interpretation to solve the satisfiability
problem.  \cite{dhk2013-popl} provide a generalisation of
CDCL algorithm called, {\em Abstract Conflict Driven Clause Learning}
(ACDCL), in terms of deduction and abduction transformers,
fixed-point, interpolation and extrapolation operators on lattices.
The partial assignments in the CDCL algorithm are an abstract domain,
unit rule is the best abstract transformer, Boolean Constraint
Propagation (BCP) computes a greatest fixed-point over the abstract
domain, decision is dual widening and clause learning can be viewed as
synthesising an abstract transformer for negation.  
%Thus, ACDCL lifts learning techniques in SAT solver to operate on a non-distributive lattice -- thus equipping the analysis with negation. 
This enables the
analyser to refine the analysis and prevent enumeration behaviour.

ACDCL has so far been instantiated as a floating-point decision
procedure~\cite{DBLP:journals/fmsd/BrainDGHK14} and a program analyser
that only uses interval constraints~\cite{tacas12}.  Hence, the
analysis is limited to bounds checking in simple numerical programs.
However, verification of certain class of programs require more
expressive domains like octagons or equalities or polyhedra.  These
domains capture interesting properties of programs which intervals can
not express.  In this paper, we identify specific properties of the
CDCL algorithm that are necessary to lift the propositional skeleton
of CDCL to relational abstract domains such as template polyhedra
domains~\cite{sriram} (octagons, zones and equalities).
To this end, we present a characterisation of abstract model search
procedure and conflict analysis procedure that lift the abstract
deduction transformer and the conflict analysis procedure to
relational lattice structures.

>>>>

\section{Introduction}

%%%%%%%%%%%%%%%%%%%%%%% AI %%%%%%%%%%%%%%%%%%%%%%%%
%
Static program analysis based on abstract interpretation~\cite{CC77} has
been widely used to verify certain classes of properties for safety-critical
systems.  A standard static analyses~\cite{se2011} for program safety aims
to compute invariants as fixed-points of abstract transformers.  Most
lattice structure used in static analysis have meet and join operations. 
The meet operation precisely model conjunction, and the join operation
over-approximate disjunction.  Thus, the precision loss in static analysers
is often due to imprecise join operations.  The loss in precision is
overcome by using richer abstract domains or equipping the analysis with
disjunction -- which is often very expensive since the analysis exhibit case
enumeration behaviour.
%
%%%%%%%%%%%%%%%%%%%%%%% ACDCL %%%%%%%%%%%%%%%%%%%%%
%
On the other hand, the combination of precise conjunction in a partial
assignments domain with learning allows Conflict Driven Clause Learning
(CDCL)~\cite{cdcl} based satisfiability (SAT) solvers to reason about
disjunction without enumerating cases.  Silva et al.~\cite{tacas12, sas12,
dhk2013-popl} presented a formal recipe for applying abstract interpretation
to solve the satisfiability problem.  \cite{dhk2013-popl} provide a
generalisation of CDCL algorithm called, {\em Abstract Conflict Driven
Clause Learning} (ACDCL), in terms of deduction and abduction transformers,
fixed-point, interpolation and extrapolation operators on lattices.  The
partial assignments in the CDCL algorithm are an abstract domain, unit rule
is the best abstract transformer, Boolean Constraint Propagation (BCP)
computes a greatest fixed-point over the abstract domain, decision is dual
widening and clause learning can be viewed as synthesising an abstract
transformer for negation.
%
%Thus, ACDCL lifts learning techniques in SAT solver to operate on a non-distributive lattice -- thus equipping the analysis with negation. 
This enables the
analyser to refine the analysis and prevent enumeration behaviour.

ACDCL has so far been instantiated as a floating-point decision
procedure~\cite{DBLP:journals/fmsd/BrainDGHK14} and a program analyser that
only uses interval constraints~\cite{tacas12}.  Hence, the analysis is
limited to bounds checking in simple numerical programs.  However,
verification of certain class of programs require more expressive domains
like octagons or equalities or polyhedra.  These domains capture interesting
properties of programs which intervals can not express.  In this paper, we
identify specific properties of the CDCL algorithm that are necessary to
lift the propositional skeleton of CDCL to relational abstract domains such
as template polyhedra domains~\cite{sriram} (octagons, zones and
equalities).  
\Omit {
To this end, we present a characterisation of abstract model
search procedure and conflict analysis procedure that lift the abstract
deduction transformer and the conflict analysis procedure to relational
lattice structures.
}
